### 2019-12-03 Mathematical Statistics/04_sampling_distributions.rmd
mathematical statistics, ch 4
mathematical statistics, ch 4

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Sampling Distributions}
   \rfoot{\color{headergrey}Chapter 4}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

@@

@@

>>>>

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Sampling Distributions}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 4}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

@@

$\sigma = (10 + 15) / sqrt( 10 + 15 - 1) = 3.1$

@@

### 4.20

Let $X_1, X_2, \ldots, X_9 \sim^{i.i.d.} N(7, 3^2)$, and $Y_1, Y_2, \ldots, X_{12} \sim^{i.i.d.} N(10, 5^2)$.

Let $W = \bar{X} - \bar{Y}$.

a.) Give the sampling distribution of $W$.

$\sigma = (3 + 5) / sqrt( 9 + 12 - 1 ) = 1.79$

$W = N(-3, 1.79^2)$

b.) Simulate the sampling distribution of W in R, and plot your results.

```{r, echo = T, fig.height=3.5, fig.width=8}
N <- 10e3
result <- numeric(N)

for(i in 1:N)
{
   X <- rnorm(9, 7, 3)
   Y <- rnorm(12, 10, 5)
   result[i] <- mean(X) - mean(Y)
}

ggplot(data.table(x = result)) +
   geom_histogram(aes(x, fill = ..count..), bins = 30) +
   geom_vline(xintercept = -3, col = "darkorange", lwd = 1.5) +
   scale_y_continuous(labels = comma) +
   labs(title = "Sampling Distribution")
```

c.) Use your simulation to find $P(W < -1.5)$.

```{r, echo = T}
phat <- mean(result <= -1.5)

p <- pnorm(-1.5, -3, 1.79)
```

Calculate an exact answer and compare.

$P(W < 1.5) = `r round(p, 4)*100`$%

### 4.21

Let $X_1, X_2, \ldots, X_N$ be a random sample from $N(0, 1)$. Let $W = X^2_1 + X^2_2 + \ldots + X^2_n$

What is the mean and variance of the sampling distribution of W?

$\mu = 0$, $\sigma = 1$

Repeat using N = 4, N = 5.

$N = 4, \sigma = 4 / sqrt(4 - 1) = 2.3$

$N = 5, \sigma = 5 / sqrt(5 - 1) = 2.5$

What observations or conjectures do you have for general __n__?

### 4.22

Let $X$ be a uniform random variable on the interval $[40, 60]$ and Y be a uniform random variable on $[45, 80]$.

Assume that X and Y are independent.

a.) Compute the expected value and variance of $X + Y$.

$\mu = 112.5$, $Var = 1/24*(140 - 85)^2 = 126.04$

b.) Simulate a sampling distribution of $X + Y$.

```{r, echo = T, fig.height=3.5, fig.width=8}
X <- runif(1000, 40, 60)
Y <- runif(1000, 45, 80)

total <- X + Y

ggplot(data.table(value = total)) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs("X + Y Sampling Distribution")

Describe the sampling distribution of $X + Y$. __Approximately Normal__

Compute the mean and variance of the sampling distribution and compare this with the theoretical mean and variance.

```{r, echo = T}
mu <- mean(total)
var <- var(total)
```

$\mu = `r mu`$, $Var = `r var`$

c.) Suppose the time (in minutes) Andy takes to complete his statistics homework is $Unif[40, 60]$ and the time Adam takes is $Unif[45, 80]$.

Assume they work independently.

One day they announce that their total time to finish an assignment was less than 90 minutes.

How likely is this?

```{r, echo = T, fig.height=3.5, fig.width=8}
p <- punif(90, 85, 140)
```

### 4.23

Let $X_1, X_2, \ldots, X_{20} \sim^{i.i.d.} Exp(2)$. Let $X = \sum^{20}_{i=1}X_i$.

a.) Simulate the sampling distribution of $X$ in R.

for(i in 1:N)
{
   samp <- rexp(n = 20, rate = 2)
   result[i] <- sum(samp)
}

b.) From your simulation, find $\mathbb{E}[X]$ and $Var[X]$.

```{r, echo = T}
mu <- mean(result)
var <- var(result)
```

$\mu = `r mu`$, $Var = `r var`$

c.) From your simulation, find $P( X \le 10)$.

```{r, echo = T}
phat <- mean(result <= 10)
```

Probablity: _`r round(phat, 4)*100`%_

### 4.24

Let $X_1, X_2, \ldots, X_{30} \sim^{i.i.d.} Exp(1/3)$ and let $\bar{X}$ denote the sample mean.

a.) Simulate the sampling distribution of $\bar{X}$ in R.

```{r, echo = T, fig.height=3.5, fig.width=8}
N <- 10e3
result <- numeric(N)

for(i in 1:N)
{
   samp <- rexp(n = 30, rate = 1/3)
   result[i] <- mean(samp)
}
```

b.) Find the mean and standard error of the sampling distribution, and compare with the theoretical results.

Sample: $\mu = `r mu`$, $\sigma = `r sigma`$

Theoretical: $\mu = \frac{\lambda}{1} = 3$, $sigma = 3/sqrt(30 - 1) = .56$

c.) From your simulation, find $P(\bar{X} \le 3.5)$.

```{r, echo = T}
phat <- mean(result <= 3.5)
```

$\hat{p} = `r phat`$

d.) Estimate $P(\bar{X} \le 3.5)$ by assuming the CLT approximation holds.

Compare this result with the one in part __(c)__.

z <- (3.5 - 3) / sigma
p <- pnorm(z)
```

$p = `r round(p, 4)`$

### 4.25

Consider the exponential distribution with density $f(x) \frac{1}{20}e^{-x/20}$, with mean and standard deviation of 20.

a.) Calculate the median of this distribution.

$0.5 = \int_{m}^{\inf}f(x)dx = -e^{-x/20}$

$\ldots = M = 20*log(2)$

$\ldots = `r 20*log(2)`$

b.) Using R, draw a random sample of size 50 and graph the histogram.

```{r, echo = T, fig.height=3.5, fig.width=8}
samp <- rexp( n = 50, rate = 1/20)

ggplot(data.table(x = samp)) +
   geom_histogram(aes(x, fill = ..count..)) +
   labs(title = "RExp(50, 1/20)")
```

What are the mean and standard deviation of your sample?

```{r, echo = T}
mu <- mean(samp)
sigma <- sd(samp)
```

c.) Run a simulation to find the (approximate) sampling distribution for the median of sample size 50 from the exponential distribution and describe it.

```{r, echo = T, fig.height=3.5, fig.width=8}
N <- 10e3
result <- numeric(N)

for(i in 1:N)
{
   samp <- rexp( n = 50, rate = 1/20)
   result[i] <- median(samp)
}

ggplot(data.table(x = result)) +
   geom_histogram(aes(x, fill = ..count..), bins = 30) +
   scale_y_continuous(labels = comma) +
   labs(title = "Sampling Distribution")
```

What is the mean and the standard error of this sampling distribution?

```{r, echo = T}
mu <- mean(result)
sigma <- sd(result) / sqrt(50)
```

d.) Repeat the above but use sample sizes $n = 100, 500 \ and \ 1,000$.

```{r, echo = T, fig.height=3, fig.width=8}
N <- 10e3
result <- numeric(N)

for(i in 1:N)
{
   samp <- rexp( n = 100, rate = 1/20)
   result[i] <- median(samp)
}

ggplot(data.table(x = result)) +
   geom_histogram(aes(x, fill = ..count..), bins = 30) +
   scale_y_continuous(labels = comma) +
   labs(title = "Sampling Distribution")
```

```{r, echo = T}
mu <- mean(result)
sigma <- sd(result) / sqrt(50)
```

```{r, echo = T, fig.height=3, fig.width=8}
N <- 10e3
result <- numeric(N)

for(i in 1:N)
{
   samp <- rexp( n = 500, rate = 1/20)
   result[i] <- median(samp)
}

ggplot(data.table(x = result)) +
   geom_histogram(aes(x, fill = ..count..), bins = 30) +
   scale_y_continuous(labels = comma) +
   labs(title = "Sampling Distribution")
```

```{r, echo = T}
mu <- mean(result)
sigma <- sd(result) / sqrt(50)
```

```{r, echo = T, fig.height=3, fig.width=8}
N <- 10e3
result <- numeric(N)

for(i in 1:N)
{
   samp <- rexp( n = 1000, rate = 1/20)
   result[i] <- median(samp)
}

ggplot(data.table(x = result)) +
   geom_histogram(aes(x, fill = ..count..), bins = 30) +
   scale_y_continuous(labels = comma) +
   labs(title = "Sampling Distribution")
```

```{r, echo = T}
mu <- mean(result)
sigma <- sd(result) / sqrt(50)
```

How does sample size affect the sampling distribution?

_The sample mean and standard error converge to the analytical solution with increased sample size._

### 4.26

Prove theorem 4.2.1.

### 4.27

Let $X_1, X_2 \sim^{i.i.d.} F$ with corresponding pdf $f(x) = \frac{2}{x^2}$, $1 \le x \le 2$.

a.) Find the pdf of $X_{max}$.

$F_{max}(x) = 8(\frac{1}{x^2} - \frac{1}{x^3})$

b.) Find the expected value of $X_{max}$.

$\mathbb{E}{[F_{max}]} = 1.545$
